---
title: ""
author: "Authors: "
date: "25/12/2025"
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    theme: cosmo
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

#0. Load packages and data wrangle

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
pacman::p_load(
  tidyverse, ggplot2, cowplot, patchwork, ggpubr, lme4, emmeans, 
  MCMCglmm, nadiv, broom.mixed, lmerTest, ggeffects, ggforce, performance,
  car
)
library("dotwhisker")  ## coefficient plots
library(forcats) #Reverse plot order
```

Load & wrangle data
```{r}
# Feather structure data
allMeasur_db <- read.csv("../RawData/bbdd_barbulas_fin_28.09.25.csv", sep = ",", header = TRUE)

##There were some ids in the feathers that were not properly written change those
#PB09->PB9
#PB08->PB8
#MI09->MI9
#2AZ11188->2A211188

allMeasur_db <- allMeasur_db %>%
  mutate(id = case_when(
    id == "PB09"     ~ "PB9",
    id == "PB08"     ~ "PB8",
    id == "MI09"     ~ "MI9",
    id == "2AZ11188" ~ "2A211188",
    TRUE             ~ id  
  ))

# Reduce the dataset to get one observation per feather a bit in line with Phillomore's comment in JAE (using the barbules independently may be too much)

MeanMeasur_db <- allMeasur_db %>%
  group_by(im_name) %>%
  summarise(
    id = first(id),
    mean_density = mean(barbule_density, na.rm = TRUE),
    .groups = "drop"
  )

# Load the colouration dataset to include relevant data
ColourInfo_db <- read.csv("../RawData/EU_GTcolour.csv", sep = ",", header = TRUE)

MeanColour_db <- ColourInfo_db %>%
  group_by(id) %>%
  summarise(
    samp_date = first(samp_date),
    habitat  = first(habitat),
    area     = first(area),
    sex      = first(sex),
    age      = first(age),
    .groups = "drop")

# Combine feather structure data bases
FeatherAd_db <- MeanColour_db %>%
  left_join(MeanMeasur_db, by = "id")

summary(FeatherAd_db)

rm(allMeasur_db, ColourInfo_db, MeanColour_db, MeanMeasur_db)

#add elevation
elevation <- read.csv("C:/Users/Usuario/Dropbox/3.CEFE/10.Color con Pablo/Colouration & Cities/DataBase_Structure/Analyses/Eobs for salmon/elevation.csv")

elevation2<-elevation %>% 
  mutate(habitat=area,
         area=city)

str(elevation2)

elevation2$habitat <- ifelse(elevation2$habitat == "Rural", "Forest", elevation2$habitat)
elevation2$area <- ifelse(elevation2$area == "Gothenburg", "Goteborg", elevation2$area)

elevation2<-elevation2 %>% 
  select(habitat, area, elevation)

FeatherAd_db<-left_join(FeatherAd_db, elevation2, by=c("habitat", "area"))



# # Combine feather info and environmental
# Temperature_db <- read.csv("../RawData/temperature_data_Fstructure.csv", sep = ",", header = TRUE)
# 
# FeatherAd_db <- FeatherAd_db %>%
#   mutate(area_habitat = paste(area, habitat, sep = "_")) # link variable for the join
# 
# Temperature_db <- Temperature_db %>%
#   mutate(area_habitat = paste(area, habitat, sep = "_")) # link variable for the join
# 
# FeatherAd_db <- FeatherAd_db %>%
#   left_join(
#     Temperature_db %>%
#       select(-X, -area, -habitat),
#     by = "area_habitat")
# 
# FeatherAd_db$sex <- factor(FeatherAd_db$sex)
# FeatherAd_db$habitat <- factor(FeatherAd_db$habitat)
# FeatherAd_db <- FeatherAd_db %>%
#   filter(habitat != "#N/A")
# FeatherAd_db$population <- paste(FeatherAd_db$area, FeatherAd_db$habitat)
# 
# FeatherAd_db <- FeatherAd_db %>%
#   mutate(season = dplyr::recode(area,
#     "Goteborg" = "breeding",
#     "Lisbon" = "wintering",
#     "Madrid" = "breeding",
#     "Milan" = "wintering",
#     "Malmo" = "breeding"
#   ))
# 
# str(FeatherAd_db)

### Prepare a PCA with feather structure
#-No need for this, we only have one variable that is density
# feather_matrix <- FeatherAd_db %>%
#   dplyr::select(mean_length, mean_barbule, mean_density)
# 
# feather_matrix <- feather_matrix %>%
#   na.omit()
# 
# feather_pca <- prcomp(feather_matrix, scale. = TRUE)
# summary(feather_pca)
# feather_pca$rotation

####Plotting stuff###
#For plotting title as a grid (https://github.com/tidyverse/ggplot2/issues/2344)
# element_textbox <- function(...) {
#   el <- element_text(...)
#   class(el) <- c("element_textbox", class(el))
#   el
# }
# 
# element_grob.element_textbox <- function(element, ...) {
#   text_grob <- NextMethod()
#   rect_grob <- element_grob(calc_element("strip.background", theme_bw()))
#   
#   ggplot2:::absoluteGrob(
#     grid::gList(
#       element_grob(calc_element("strip.background", theme_bw())),
#       text_grob
#     ),
#     height = grid::grobHeight(text_grob), 
#     width = grid::unit(1, "npc")
#   )
# }
# 
# #####Sumamry of sample sizes####
# TableSummary_adults <- FeatherAd_db %>%
#   # filter(YH < 500) %>%
#   group_by(habitat, sex, age) %>%
#   # select(YB, YH, YC, habitat, sex, age, id) %>%
#   select(mean_YB, habitat, sex, age, id) %>%
#   count (id)  %>%
#   summarise_all(funs(Nind = n()))
# 
# #Sumamry of sample sizes
# TableSummary_MALMO <- MALMO %>%
#   filter(YH < 500) %>% #Remove observations over 500
#   group_by(habitat, sex, age) %>%
#   # select(YB, YH, YC, habitat, sex, age, id) %>%
#   select(YC, habitat, sex, age, id) %>%
#   count (id)  %>%
#   summarise_all(funs(Nind = n()))
# library(patchwork)
# AdultYH + NestlingsYH

```

#1. Are there differences in the feather microstructure between the urban and rural habitat
##1.1 Model
```{r}
FeatherAd_db$area_habitat<-paste(FeatherAd_db$area, FeatherAd_db$habitat, sep="_")

model_habitat <- lmer(mean_density ~ habitat + sex + age + habitat*age + (1|area_habitat)+ (1|area) +(1|id), data=FeatherAd_db) 
summary(model_habitat)
anova(model_habitat)

plot(predict_response(model_habitat, terms="elevation"))

#some tests
hist(residuals(model_habitat))
qqnorm(resid(model_habitat))
qqline(resid(model_habitat), col = "red")

# Fixed effects:
#                      Estimate Std. Error       df t value Pr(>|t|)    
# (Intercept)           19.7953     0.4390  26.1403  45.092   <2e-16 ***
# habitatUrban          -0.1566     0.3116  17.4651  -0.502   0.6216    
# sex                    0.3267     0.1895 197.6403   1.724   0.0863 .  
# age2yr+                0.4494     0.2631 200.3492   1.708   0.0892 .  
# habitatUrban:age2yr+  -0.3476     0.3744 199.8574  -0.929   0.3543      

model_habitat_2 <- lmer(mean_density ~ habitat + sex + age + (1|area_habitat)+(1|area) +(1|id), data=FeatherAd_db) 
summary(model_habitat_2)
anova(model_habitat_2)

# Random effects:
#  Groups       Name        Variance Std.Dev.
#  id           (Intercept) 1.17968  1.0861  
#  area_habitat (Intercept) 0.04145  0.2036  
#  area         (Intercept) 0.28503  0.5339  
#  Residual                 1.48197  1.2174  
# Number of obs: 624, groups:  id, 208; area_habitat, 10; area, 5
# 
# Fixed effects:
#              Estimate Std. Error       df t value Pr(>|t|)    
# (Intercept)   19.8645     0.4284  26.3436  46.366   <2e-16 ***
# habitatUrban  -0.3517     0.2293   5.3056  -1.534   0.1824    
# sex            0.3311     0.1894 198.7088   1.748   0.0821 .  
# age2yr+        0.2839     0.1940 199.6841   1.464   0.1449    


#some tests
hist(residuals(model_habitat_2))
qqnorm(resid(model_habitat_2))
qqline(resid(model_habitat_2), col = "red")
```

No hay una interaccion significativa entre habitat y age; testeado en relacion a los resultados de Sandor (2022). Quito la interaccion y en el modelo simple no hay efectos (generales) de lo urbano. Me ha llamado la atencion la cantidad de varianza que explica el individuo (37.5%).

##1.2 Plot
```{r}
library(ggeffects)
library(ggplot2)

pred_general <- as.data.frame(predict_response(model_habitat, terms = "habitat"))

p1 <- ggplot() +
  # --- Jittered raw data points ---
  geom_jitter(
    data = FeatherAd_db,
    aes(x = habitat, y = mean_density),
    width  = 0.15,
    height = 0,
    size   = 1.8,
    alpha  = 0.35,
    colour = "grey70",
    shape  = 19
  ) +
    # --- Confidence interval error bars ---
  geom_errorbar(
    data = pred_general,
    aes(x = x, ymin = predicted-std.error, ymax = predicted+std.error, colour = x),
    width     = 0.08,
    linewidth = 1.0
  ) +
  # --- Predicted mean points ---
  geom_point(
    data = pred_general,
    aes(x = x, y = predicted, fill = x),
    size  = 5,
    shape = 21,
    stroke = 1.5
  ) +

  # --- Colours for Forest vs Urban ---
  scale_fill_manual(
    values = c(
      "Forest" = "springgreen4",
      "Urban"  = "grey50"
    )
  ) +
    scale_color_manual(
    values = c(
      "Forest" = "black",
      "Urban"  = "black"
    )
  ) +

  labs(
    x      = "Habitat",
    y      = "Mean barbule density"
  ) +

  theme_bw(base_size = 22) +
  theme(
    legend.position = "none",
    axis.text.x     = element_text(size = 12, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.title.x    = element_blank(),
  )+
  scale_y_continuous(
  limits = c(14.9, 27),           # min and max of y axis
  breaks = seq(16, 26, by = 2)  # breaks every 2 units: 16, 18, 20, 22, 24, 26
)
  
p1
```

#2. Are the urban effects consisten across space
##2.1 Model
```{r}
model_space <- lmer(mean_density ~ habitat*area + sex + age + (1|id), data=FeatherAd_db) 
summary(model_space)
anova(model_space) 

#the interaction is not significant, meaning the (lack of) effects of habitat are consistent in space
model_space_2 <- lmer(mean_density ~ habitat+area + sex + age + (1|id), data=FeatherAd_db) 
summary(model_space_2)
anova(model_space_2) 

```

Post-hoc test
```{r}
library(emmeans)

# --- Marginal means for area (averaged over habitat, sex, age) ---
emm_area <- emmeans(model_space_2, ~ area)
emm_area  # view the marginal means + SE

# --- All pairwise comparisons ---
pairs_area <- pairs(emm_area, adjust = "tukey")  # or "bonferroni", "fdr", "none"
summary(pairs_area)

# --- If you want confidence intervals too ---
confint(pairs_area)

# --- Compact letter display (groups sharing a letter are not significantly different) ---
library(multcomp)
cld(emm_area, adjust = "tukey", Letters = letters)

# --- Extract CLD letters ---
cld_df <- as.data.frame(cld(emm_area, adjust = "tukey", Letters = letters))
cld_df$.group <- trimws(cld_df$.group)  # remove whitespace from letters
```

##2.2 Plot
```{r}
# --- Model-adjusted predictions by area, coloured by habitat ---
pred_space <- predict_response(model_space, terms = c("area", "habitat"))
pred_space_df <- as.data.frame(pred_space)

# --- Calculate label position (just above the upper CI of the highest group per area) ---
label_pos <- pred_space_df %>%
  group_by(x) %>%
  summarise(label_y = max(predicted + std.error) + 0.3)  # adjust 0.3 to taste

cld_df <- cld_df %>%
  left_join(label_pos, by = c("area" = "x"))

# --- Summarise raw data for jitter (unadjusted, just for visual context) ---
raw_summary <- FeatherAd_db %>%
  group_by(area, habitat) %>%
  summarise(mean_density = mean(mean_density), .groups = "drop")

# --- Colour palette ---
hab_colours <- c("Forest" = "springgreen4", "Urban" = "grey50")

p2 <- ggplot() +
  # --- Raw jittered points ---
  geom_jitter(
    data = FeatherAd_db,
    aes(x = area, y = mean_density, colour = habitat),
    width  = 0.2,
    height = 0,
    size   = 1.5,
    alpha  = 0.3,
    shape  = 16
  ) +
  # --- SE error bars ---
  geom_errorbar(
    data = pred_space_df,
    aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error,
        colour = group),
    width     = 0.1,
    linewidth = 0.9,
    position  = position_dodge(width = 0.5)
  ) +
  # --- Predicted means ---
  geom_point(
    data = pred_space_df,
    aes(x = x, y = predicted, fill = group),
    size     = 4,
    shape    = 21,
    stroke   = 1.2,
    position = position_dodge(width = 0.5)
  ) +
  # --- CLD letters at fixed y = 16 ---
  geom_text(
    data     = cld_df,
    aes(x    = area, label = .group),
    y        = 14.7,
    size     = 4,
    colour   = "black"
  ) +
  scale_y_continuous(
    limits = c(15.5, NA),              # must be below y = 16 so letters are visible
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  scale_fill_manual(values = hab_colours) +
  scale_colour_manual(values = hab_colours) +
  labs(
    y      = "",
    fill   = "Habitat",
    colour = "Habitat"
  ) +
  theme_bw(base_size = 22) +
  theme(
    legend.position = "none",
    axis.title.x    = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )+
  scale_y_continuous(
  limits = c(14.9, 27),           # min and max of y axis
  breaks = seq(16, 26, by = 2)  # breaks every 2 units: 16, 18, 20, 22, 24, 26
)

p2
```

#3. Links with latitude
##3.1 Model latitude
```{r}
city_lat <- c(
  Goteborg = 57.7089,
  Lisbon   = 38.7169,
  Madrid   = 40.4168,
  Malmo    = 55.6050,
  Milan    = 45.4642
)

city_elev <- c(
  Goteborg = 12,
  Lisbon   = 2,
  Madrid   = 667,
  Malmo    = 17,
  Milan    = 122
)

FeatherAd_db$latitude  <- city_lat[FeatherAd_db$area]
FeatherAd_db$elevation <- city_elev[FeatherAd_db$area]


model_latitude<- lmer(mean_density ~ habitat+latitude + sex + age + (1|id)+(1|area), data=FeatherAd_db) 
summary(model_latitude)
anova(model_latitude) 

plot(predict_response(model_latitude, terms=c("latitude", "habitat")))


model_elevation<- lmer(mean_density ~ habitat+elevation + sex + age + (1|id)+(1|area), data=FeatherAd_db) 
summary(model_elevation)
anova(model_elevation) 

plot(predict_response(model_elevation, terms=c("elevation", "habitat")))
```

#4. Euro Map
```{r}
library(maps)

# Your data
city_lat <- c(Goteborg = 57.7089, Lisbon = 38.7169, Madrid = 40.4168, Malmo = 55.6050, Milan = 45.4642)
city_lon <- c(Goteborg = 11.9746, Lisbon = -9.1395, Madrid = -3.7038, Malmo = 13.0038, Milan = 9.1900)

cities_df <- data.frame(
  city = names(city_lat),
  lat  = city_lat,
  lon  = city_lon
)

# Get all world countries and crop to the range of the points
world <- map_data("world")

p_map<-ggplot() +
  geom_polygon(data = world, aes(x = long, y = lat, group = group),
               fill = "lightgrey", color = "white", linewidth = 0.2) +
  geom_point(data = cities_df, aes(x = lon, y = lat),
             color = "firebrick", size = 3) +
  geom_text(data = cities_df, aes(x = lon, y = lat, label = city),
            vjust = -0.8, size = 3.5) +
  coord_fixed(1.3, xlim = c(-12, 20), ylim = c(35, 62)) +
  labs(x = "Longitude", y = "Latitude") +
  theme_bw()
```

```{r}
library(cowplot)

plot<-plot_grid(p1, p2,
          ncol    = 2,
          align   = "h",
          rel_widths = c(1, 2))

#save it
ggsave(plot,
  filename = "plot_effects.png",
  width    = 12,
  height   = 6,
  dpi      = 300
)
```














#Analyses
##Linking colour and structure
#add latitude
```{r}
city_lat <- c(
  Goteborg = 57.7089,
  Lisbon   = 38.7169,
  Madrid   = 40.4168,
  Malmo    = 55.6050,
  Milan    = 45.4642
)


city_elev <- c(
  Goteborg = 12,
  Lisbon   = 2,
  Madrid   = 667,
  Malmo    = 17,
  Milan    = 122
)

FeatherAd_db$latitude  <- city_lat[FeatherAd_db$area]
FeatherAd_db$elevation <- city_elev[FeatherAd_db$area]

table(FeatherAd_db$age)
```

Me understandind the dataset
```{r}
table(FeatherAd_db$area)
table(FeatherAd_db$habitat)
table(FeatherAd_db$season)
table(FeatherAd_db$age)
table(FeatherAd_db$id)
table(FeatherAd_db$samp_date)

hist(FeatherAd_db$mean_density)
```

#feathers were samples in two different years but this not appears in the data - create these variables
```{r}
str(FeatherAd_db)
FeatherAd_db<-FeatherAd_db %>% 
  mutate(year = substr(samp_date, 7, nchar(samp_date))) %>% 
  mutate(year = case_when(
    year == "14" ~ "2014",
    year == "15" ~ "2015",
    year == "18" ~ "2014",
    year == "24" ~ "2014",
    TRUE         ~ year
  ))

table(FeatherAd_db$year)

#just realised that the samples in 2015 are only present in one area
```


#Test whether there are differences between urban/forest
```{r}
model_habitat <- lmer(mean_density ~ habitat+latitude +elevation  + sex + age + elevation + (1|area), data=FeatherAd_db) 
summary(model_habitat)
anova(model_habitat)

hist(residuals(model_habitat))


plot(predict_response(model_habitat, terms="habitat"), show_data = TRUE)
```

#Urban vs Forest
```{r}
library(ggeffects)
library(ggplot2)

pred <- as.data.frame(predict_response(model_habitat, terms = "habitat"))


p1 <- ggplot() +
  # --- Jittered raw data points ---
  geom_jitter(
    data = FeatherAd_db,
    aes(x = habitat, y = mean_density),
    width  = 0.15,
    height = 0,
    size   = 1.8,
    alpha  = 0.35,
    colour = "grey70",
    shape  = 19
  ) +
    # --- Confidence interval error bars ---
  geom_errorbar(
    data = pred,
    aes(x = x, ymin = predicted-std.error, ymax = predicted+std.error, colour = x),
    width     = 0.08,
    linewidth = 1.0
  ) +
  # --- Predicted mean points ---
  geom_point(
    data = pred,
    aes(x = x, y = predicted, fill = x),
    size  = 5,
    shape = 21,
    stroke = 1.5
  ) +

  # --- Colours for Forest vs Urban ---
  scale_fill_manual(
    values = c(
      "Forest" = "springgreen4",
      "Urban"  = "grey50"
    )
  ) +
    scale_color_manual(
    values = c(
      "Forest" = "black",
      "Urban"  = "black"
    )
  ) +

  labs(
    x      = "Habitat",
    y      = "Predicted barbule density"
  ) +

  theme_bw(base_size = 18) +
  theme(
    legend.position = "none",
    axis.text.x     = element_text(size = 12, face = "bold"),
    plot.title      = element_text(hjust = 0.5, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.title.x    = element_blank(),
  )+
  scale_y_continuous(
  limits = c(16, 24.4),           # min and max of y axis
  breaks = seq(16, 24, by = 2)  # breaks every 2 units: 16, 18, 20, 22, 24, 26
)
  
p1



# Warning message:
# Removed 4 rows containing missing values or values outside the scale range (`geom_point()`). 
## There are 4 NAs

###Tengo que elegir los colores
```

#Test whether there are differences between areas and habitats
```{r}
FeatherAd_db$area<-as.factor(FeatherAd_db$area)
FeatherAd_db$habitat<-as.factor(FeatherAd_db$habitat)
FeatherAd_db$age<-as.factor(FeatherAd_db$age)
FeatherAd_db$sex<-as.factor(FeatherAd_db$sex)

model_area <- lm(mean_density ~ area+habitat+age + sex + age, data=FeatherAd_db) 
summary(model_area)
anova(model_area) #there are differences between areas

pred_means <- predict_response(model_area, terms = c("area"))
as.data.frame(pred_means)


plot(predict_response(model_area, terms=c("area")), show_data = FALSE) #they do not seem huge

#some rejigging to plot them latitudinally

# Set area as factor ordered by latitude (low to high)
FeatherAd_db2<-FeatherAd_db
FeatherAd_db2$area <- factor(FeatherAd_db2$area, 
                             levels = c("Lisbon", "Madrid", "Milan", "Malmo", "Goteborg"))

# Refit the model with the reordered factor
model_area_habitat <- lmer(mean_density ~ habitat + elevation + sex + age + (1|area), data = FeatherAd_db2)
summary(model_area_habitat)
anova(model_area_habitat)

plot(predict_response(model_area_habitat, terms="elevation"), show_data = TRUE)

# Plot
plot(predict_response(model_area, terms = c("area", "age")), show_data = TRUE)

# sex differences
model_sex_age <- lmer(mean_density ~ habitat + sex + age + (1|area), data = FeatherAd_db2)
summary(model_sex_age)

plot(predict_response(model_sex_age, terms = c("sex")), show_data = FALSE)
plot(predict_response(model_sex_age, terms = c("age")), show_data = FALSE)

#effects of latitude
model_latitude <- lmer(mean_density ~ habitat + sex + age + latitude + (1|area), data = FeatherAd_db2)
summary(model_latitude)

#difernecia
str(FeatherAd_db2)

result <- FeatherAd_db2 %>%
  group_by(area, habitat) %>%
  summarise(
    mean_density = mean(mean_density, na.rm = TRUE),
    mean_lat = mean(latitude, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = habitat, values_from = mean_density) %>%
  mutate(density_diff = Urban - Forest)

model_cosa <- lm(density_diff ~ mean_lat, data = result)
summary(model_cosa)

```

#By city
```{r}
# library(ggeffects)
# library(ggplot2)
# 
# pred_city_habitat <- as.data.frame(predict_response(model_area_habitat, terms = c("habitat", "area")))
# pred_city_habitat$com<-paste(pred_city_habitat$group, pred_city_habitat$x, sep="_")
# 
# p1 <- ggplot() +
#   # --- Jittered raw data points ---
#   geom_jitter(
#     data = FeatherAd_db2,
#     aes(x = area_habitat, y = mean_density),
#     width  = 0.15,
#     height = 0,
#     size   = 1.8,
#     alpha  = 0.35,
#     colour = "grey70",
#     shape  = 19
#   ) +
#     # --- Confidence interval error bars ---
#   geom_errorbar(
#     data = pred_city_habitat,
#     aes(x = x, ymin = predicted-std.error, ymax = predicted+std.error, colour = x),
#     width     = 0.08,
#     linewidth = 1.0
#   ) +
#   # --- Predicted mean points ---
#   geom_point(
#     data = pred_city_habitat,
#     aes(x = x, y = predicted, fill = x),
#     size  = 5,
#     shape = 21,
#     stroke = 1.5
#   ) +
# 
#   # --- Colours for Forest vs Urban ---
#   scale_fill_manual(
#     values = c(
#       "Forest" = "springgreen4",
#       "Urban"  = "grey50"
#     )
#   ) +
#     scale_color_manual(
#     values = c(
#       "Forest" = "black",
#       "Urban"  = "black"
#     )
#   ) +
# 
#   labs(
#     x      = "Habitat",
#     y      = "Predicted barbule density"
#   ) +
# 
#   theme_bw(base_size = 18) +
#   theme(
#     legend.position = "none",
#     axis.text.x     = element_text(size = 12, face = "bold"),
#     plot.title      = element_text(hjust = 0.5, face = "bold"),
#     panel.grid.major.x = element_blank(),
#     panel.grid.minor.x = element_blank()
#   )
# 
# p1

# Warning message:
# Removed 4 rows containing missing values or values outside the scale range (`geom_point()`). 
## There are 4 NAs

###Tengo que elegir los colores


library(ggeffects)
library(ggplot2)
library(dplyr)

# --- Model-adjusted predictions by area, coloured by habitat ---
pred <- predict_response(model_area_habitat, terms = c("area", "habitat"))
pred_df <- as.data.frame(pred)

# --- Summarise raw data for jitter (unadjusted, just for visual context) ---
raw_summary <- FeatherAd_db2 %>%
  group_by(area, habitat) %>%
  summarise(mean_density = mean(mean_density), .groups = "drop")

# --- Colour palette ---
hab_colours <- c("Forest" = "springgreen4", "Urban" = "grey50")

p2 <- ggplot() +
  # --- Raw jittered points ---
  geom_jitter(
    data = FeatherAd_db2,
    aes(x = area, y = mean_density, colour = habitat),
    width  = 0.2,
    height = 0,
    size   = 1.5,
    alpha  = 0.3,
    shape  = 16
  ) +
  # --- SE error bars ---
  geom_errorbar(
    data = pred_df,
    aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error,
        colour = group),
    width     = 0.1,
    linewidth = 0.9,
    position  = position_dodge(width = 0.5)
  ) +
  # --- Predicted means ---
  geom_point(
    data = pred_df,
    aes(x = x, y = predicted, fill = group),
    size     = 4,
    shape    = 21,
    stroke   = 1.2,
    position = position_dodge(width = 0.5)
  ) +
  # --- CLD letters at fixed y = 16 ---
  geom_text(
    data     = cld_df,
    aes(x    = area, label = .group),
    y        = 15.8,
    size     = 4,
    colour   = "black"
  ) +
  scale_y_continuous(
    limits = c(15.5, NA),              # must be below y = 16 so letters are visible
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  scale_fill_manual(values = hab_colours) +
  scale_colour_manual(values = hab_colours) +
  labs(
    y      = "Mean barbule density",
    fill   = "Habitat",
    colour = "Habitat"
  ) +
  theme_bw(base_size = 18) +
  theme(
    legend.position = "none",
    axis.title.x    = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )+
  scale_y_continuous(
  limits = c(16, 24.4),           # min and max of y axis
  breaks = seq(16, 24, by = 2)  # breaks every 2 units: 16, 18, 20, 22, 24, 26
)

p2
```

```{r}
library(emmeans)

# --- Marginal means for area (averaged over habitat, sex, age) ---
emm_area <- emmeans(model_area_habitat, ~ area)
emm_area  # view the marginal means + SE

# --- All pairwise comparisons ---
pairs_area <- pairs(emm_area, adjust = "tukey")  # or "bonferroni", "fdr", "none"
summary(pairs_area)

# --- If you want confidence intervals too ---
confint(pairs_area)

# --- Compact letter display (groups sharing a letter are not significantly different) ---
library(multcomp)
cld(emm_area, adjust = "tukey", Letters = letters)

# --- Extract CLD letters ---
cld_df <- as.data.frame(cld(emm_area, adjust = "tukey", Letters = letters))
cld_df$.group <- trimws(cld_df$.group)  # remove whitespace from letters

# --- Calculate label position (just above the upper CI of the highest group per area) ---
label_pos <- pred_df %>%
  group_by(x) %>%
  summarise(label_y = max(predicted + std.error) + 0.3)  # adjust 0.3 to taste

cld_df <- cld_df %>%
  left_join(label_pos, by = c("area" = "x"))

```








#De aqui pabajo son cosas de Pablo
##Adults
### Density adults (5 population pairs)
```{r Chroma ANALYSIS Adults 5 pop., echo=TRUE, message=FALSE, warning=FALSE,fig.align="center",fig.height=12, fig.width=8}
#PreAnalysis regarding season of feather collection
model <- lmer(scale(mean_density) ~ factor(habitat) + factor(sex) + factor(age) + mean_YB + (1|area) +(1|area_habitat), REML='T',data=FeatherAd_db) #Model for the stats section
plot(model)
qqnorm(resid(model))
#shapiro.test(resid(modelPre_C))
check_model(model)
# drop1( modelPre_C, test="Chisq")
summary( model)
anova( model)
#ranova(modelPre_C)
r.squaredGLMM(model)

#Repeatibilidad medidas ####
mod <- lmer(scale(YB)~sex+age+habitat
                 + (1|id),REML='T',data=subset(adults, rep == 1))#Initial Model for Chroma

mod <- lmer(scale(YC) ~ age + sex + (1 | id) + (1 | area), REML='F', data = subset(adults, habitat == "Urban"))
# # summary(VarChromaUrb)

#Bootstrapping
nsim<- 1000 #number of simulations
bsim <- sim(mod, n.sim=nsim) #bsim is the new, bootstrapped dataset

#Retrieve all fixed effect estimates (mean and 95% credible intervals) simultaneously 
colnames(bsim@fixef) = names(fixef(mod))
coef_fixef =  rbind(apply(bsim@fixef,2, mean), apply(bsim@fixef, 2, quantile, prob=c(0.025,0.975)))

# Transpose the fixed effects table
fixefTable = as.data.frame(t(as.data.frame(coef_fixef))) %>%
  tibble::rownames_to_column("coefName")

# Retrieve all random effect estimates (mean, credible intervals) simultaneously
coef_ranef = lapply(ranef(bsim), function(x) c(mean (apply(x[,,1],1,var)),  quantile(apply(x[,,1],1,var), prob=c(0.025,0.975))))
# Transpose the random effects table
ranefTable = as.data.frame(t(as.data.frame(coef_ranef))) %>%
  tibble::rownames_to_column("coefName")

# sigma is to call the residual (=within individual) variance
coef_res = c(mean(bsim@sigma^2), quantile(bsim@sigma^2, c(0.025, 0.975)))

# Transpose the residual effects table
resTable = as.data.frame(t(as.data.frame(coef_res))) %>%
  mutate(coefName = "residual")


#put together all estimates from fixed and random effects in one single Table
allCoefs= bind_rows(fixefTable, 
                          ranefTable, 
                          resTable) %>% 
  mutate(coefString = paste0(round(`V1`,2)," (",round(`2.5%`,2),", ",round(`97.5%`,2),")"))


##Repeatibility is estimated as among-individual variance / total variance
AreaUrb.Rpt = apply(bsim@ranef$area[ , , 1],1,var)/  (apply(bsim@ranef$area[ , , 1],1,var)+ apply(bsim@ranef$id[ , , 1],1,var)+bsim@sigma^2)
c(mean(AreaUrb.Rpt), quantile (AreaUrb.Rpt, prob=c(0.025,0.975)))

IDUrb.Rpt = apply(bsim@ranef$id[ , , 1],1,var)/  (apply(bsim@ranef$area[ , , 1],1,var)+ apply(bsim@ranef$id[ , , 1],1,var)+bsim@sigma^2)
c(mean(IDUrb.Rpt), quantile (IDUrb.Rpt, prob=c(0.025,0.975)))

# Urb.Resid = (bsim@sigma^2)/  (apply(bsim@ranef$area[ , , 1],1,var)+ apply(bsim@ranef$id[ , , 1],1,var)+bsim@sigma^2)
# c(mean(Urb.Resid), quantile (Urb.Resid, prob=c(0.025,0.975)))

AreaFor.Rpt = apply(bsim@ranef$area[ , , 1],1,var)/  (apply(bsim@ranef$area[ , , 1],1,var)+ apply(bsim@ranef$id[ , , 1],1,var)+bsim@sigma^2)
c(mean(AreaFor.Rpt), quantile (AreaFor.Rpt, prob=c(0.025,0.975)))

IDFor.Rpt = apply(bsim@ranef$id[ , , 1],1,var)/  (apply(bsim@ranef$area[ , , 1],1,var)+ apply(bsim@ranef$id[ , , 1],1,var)+bsim@sigma^2)
c(mean(IDFor.Rpt), quantile (IDFor.Rpt, prob=c(0.025,0.975)))

# For.Resid = (bsim@sigma^2)/ (apply(bsim@ranef$area[ , , 1],1,var)+ apply(bsim@ranef$id[ , , 1],1,var)+bsim@sigma^2)
# c(mean(For.Resid), quantile (For.Resid, prob=c(0.025,0.975)))

# AreaForest.Rpt = apply(bsim@ranef$area[ , , 1],1,var)/  (apply(bsim@ranef$area[ , , 1],1,var)+bsim@sigma^2)
# c(mean(AreaForest.Rpt), quantile (AreaForest.Rpt, prob=c(0.025,0.975)))
# 
# WithinAreaForest.Rpt = apply(bsim@ranef$area[ , , 1],1,var)/  (apply(bsim@ranef$area[ , , 1],1,var)+bsim@sigma^2)
# c(mean(AreaForest.Rpt), quantile (WithinAreaForest.Rpt, prob=c(0.025,0.975)))


### Figures ###
#Europe Map for the 5 Urban areas (Figure 1a)

```{r MAP Sampling Adults Areas, message=FALSE, warning=FALSE, include=FALSE}
#Libraries
library(tidyverse)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(sf)
library(raster)
library(ggrepel)
library(showtext)
library("png")
theme_set(theme_bw())


font_add_google(
  "Lato",
  regular.wt = 300,
  bold.wt = 700)

# read vector data using rnaturalearth
# land <- ne_download(
#   scale = 50,
#   type = "land",
#   category = "physical",
#   returnclass = "sf")
# the above doesnt work anymore because issue on the http, instead dowload the desired vector data, e.g.
# https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/50m/physical/ne_50m_land.zip
# and upload it manually with ne_load

land <- ne_load(scale = 50,
  type = "land",
  category = "physical",
  destdir = "/Users/Pablo/Downloads/ne_50m_land",
  returnclass = "sf")

countries <- ne_countries(scale='medium',returnclass = 'sf')

#urban_d<- ne_download(scale = "medium", type = "urban_areas", category = "cultural",returnclass = 'sf')
#urban<- ne_load(scale = "medium", type = "urban_areas", category = "cultural",returnclass = 'sf')

library(rgdal)
#urban_layer <- readOGR(dsn = "/Users/Pablo/Downloads/ne_50m_urban_areas", 
 #                    layer = "ne_50m_urban_areas")

# create dataframe with  locations
df <- data.frame(
  site = c("Gothenburg", "Malmö", "Milan", "Madrid", "Lisbon"),
  lat = c(57.69, 55.6,45.53,40.44,38.74),
  lon = c(11.94, 12.99,9.21,-3.73,-9.18),
  typeface=sample(c("plain","plain","plain","plain","plain"))
)


# create dataframe with  locations including meta-analyses
df <- data.frame(
  site = c("Gothenburg", "Malmö", "Milan", "Madrid", "Lisbon","Paris","Niort","Tartu","Turku","Antwerp","Barcelona"),
  lat = c(57.69, 55.6, 45.53, 40.44, 38.74, 48.85, 46.32, 58.37, 60.45, 51.22, 41.38),
  lon = c(11.94, 12.99, 9.21, -3.73, -9.18, 2.35, 0.46, 26.72, 22.26, 4.40, 2.17),
  typeface = sample(c("plain", "plain","plain","plain","plain","plain","plain","plain","plain","plain","plain")),
  freq = c(4, 1, 1, 1, 1, 1, 1, 3, 9, 4, 1)
  )

df <- data.frame(
  site = c("Gothenburg", "" , "Malmö", "Milan", "Madrid", "Lisbon","Paris","Niort","Tartu","Turku","Antwerp","Barcelona"),
  lat = c(57.69, 57.69, 55.6, 45.53, 40.44, 38.74, 48.85, 46.32, 58.37, 60.45, 51.22, 41.38),
  lon = c(11.94, 11.94, 12.99, 9.21, -3.73, -9.18, 2.35, 0.46, 26.72, 22.26, 4.40, 2.17),
  typeface = sample(c("plain", "plain", "plain","plain","plain","plain","plain","plain","plain","plain","plain","plain")),
  freq = c(3, 1, 1, 1, 1, 1, 1, 1, 3, 9, 4, 1)
  )


p <- ggplot() +

  # first layer is the land mass outline
  # as a grey border (fancy)
  geom_sf(data = land,
          fill = NA,
          color = "grey50",
          fill = "#dfdfdf",
          lwd = 1) +

  # crop the whole thing to size
  coord_sf(xlim = c(-25, 25),
           ylim = c(35, 63))

p <- p +

  # second layer are the countries
  # with a white outline and and a
  # grey fill
  geom_sf(data = countries,
          color = "white",
          fill = "#dfdfdf",
          size = 0.2)+

  # crop the whole thing to size
  coord_sf(xlim = c(-25, 25),
           ylim = c(35, 63))

p <- p +

  # add the locations of the sites
  # as a point
  geom_point(data = df, pch=21,
             aes(lon, lat, fill=site, size = freq),
             col = "grey20") +

  # use ggrepel to add fancy
  # labels nudged to a
  # longitude of -25
  geom_text_repel(
    data = df,
    aes(lon,
        lat,
        label = site,
        fontface=typeface
        ),
    nudge_x      = -25 - df$lon,
    direction    = "y",
    hjust        = 0,
    segment.size = 0.2,
    size = 5,
    seed = 1 # ensures the placing is consistent between renders
  ) +
  ylab("Latitude")+
  xlab("Longitude")+
  scale_fill_manual("site", values=c(Gothenburg = "grey93",
                              " " = "black",   
                             Malmö = "black",
                             Milan = "black",
                             Madrid = "black",
                             Lisbon = "black",
                             Paris = "grey93",
                             Niort = "grey93",
                             Tartu = "grey93",
                             Turku = "lightskyblue3",
                             Antwerp = "lightskyblue3",
                             Barcelona = "lightskyblue3"),
                    guide = "none") +
  scale_size(breaks = c(1,4,9), range = c(4, 12), labels = c("1", "4", "8 or more")) +
  theme(
                legend.position = c(0.9, 0.45),
                strip.text.x = element_text(size = 17),
                axis.ticks = element_line(colour = "black", size=1),
                axis.text.x=element_text(size=12,colour = "black"),
                axis.title.x=element_text(size=17),
                axis.text.y=element_text(size=12,colour = "black"),
                axis.title.y=element_text(size=17),
                plot.title = element_text(lineheight=2, face="bold", hjust=0.15,size=18),#
                plot.margin = unit(c(0.25,0.25,0.25,0.25), "cm"))

  # crop the whole thing to size
  coord_sf(xlim = c(-25, 25),
           ylim = c(35, 63)) 
  
p <- p + guides(size = guide_legend(title="Nr of studies"))
  
  
GT <- readPNG("/Users/Pablo/Desktop/Figures/GT_FINAL_FINAL.png")
  
# p = ggdraw() + 
#   draw_plot(p, 0, 0, 1, 1) +
#   draw_image(GT, 0.78,0.42,0.27,-0.34)

#ggsave(p, file = "/Users/Pablo/Desktop/Colour_GT/Map.pdf",
 #      width = 10, height = 10, dpi = 350, units = "in", 
  #     device="pdf")

# ggsave(p, file = "/Users/Pablo/Dropbox/Glasgow PostDoc project (EHU:UPV)/Projects/Colour_GT/Draft/Figure1_New.png",
#     width = 10, height = 10, dpi = 400, units = "in", 
#        device="png")

# ggsave(p, file = "../Figures/GraphicalAbstract/MapLocations.pdf",
#      width = 12, height = 12, dpi = 350, units = "in",
#     device="pdf")

```

#Adults Feather strucutre for the 5 Urban areas (Figure)
```{r PLOT Adults 5 pop., message=FALSE, warning=FALSE, include=FALSE}

# correlogram across structure and environemtnal variables

variables <- FeatherAd_db %>%
  select(
    mean_YC, mean_YB,
    mean_density,
    "an_mean_Tmean", "an_se_Tmean", "an_mean_Tmin", "an_se_Tmin",
    "an_mean_Tmax", "an_se_Tmax", "win_mean_Tmean", "win_se_Tmean",
    "win_mean_Tmax", "win_se_Tmax", "win_mean_Tmin", "win_se_Tmin",
    "mou_mean_Tmean", "mou_se_Tmean", "mou_mean_Tmin",
    "mou_se_Tmin", "mou_mean_Tmax", "mou_se_Tmax")

cor_matrix_structure_env <- cor(variables, use = "complete.obs")


corrplot::corrplot(cor_matrix_structure_env, method = "color", type = "upper",
         addCoef.col = "black", # show correlation coefficients
         tl.col = "black", tl.srt = 45, # axis labels
         number.cex = 0.7) 


# Summarise mean and SD per trait, area, and habitat
summary_df <- FeatherAd_db %>%
  group_by(area, habitat) %>%
  summarise(
    sd = sd(mean_density, na.rm = TRUE), 
    mean_density = mean(mean_density, na.rm = TRUE),
    .groups = "drop")

# Main plot with jittered raw points and overlaid means ± SD
FigurePopDensity <- ggplot(FeatherAd_db, aes(x = factor(habitat, levels = c("Urban", "Forest")), y = mean_density)) +
  geom_point(
    pch = 21, size = 4, fill = "grey85", colour = "grey85",
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.5),
    aes(alpha = 0.3)) +
  geom_pointrange(data = summary_df,
  aes(
    x = habitat,
    y = mean_density,           
    ymin = mean_density - sd,
    ymax = mean_density + sd,
    fill = habitat),
  colour = "black", shape = 21, size = 1,
  position = position_dodge(width = 0.5),
  inherit.aes = FALSE) +
  ylab("Barbule density") +
  xlab(" ") +
  facet_grid( ~ area, scales = "free_y") +
  scale_fill_manual(values = c("yellowgreen", "black")) +
  guides(fill = "none", alpha = "none") +
  theme_bw() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 18, hjust = 0.5, margin = margin(t = 5, b = 5)),
    strip.text = element_text(size = 17),
    axis.ticks = element_line(colour = "black", size = 1),
    axis.text.x = element_text(size = 18, colour = "black"),
    axis.title.x = element_text(size = 22),
    axis.text.y = element_text(size = 15, colour = "black"),
    axis.title.y = element_text(size = 22),
    plot.margin = unit(c(1, 1, 1, 1), "cm"))

ggsave(FigurePopDensity, file = "../Figures/FigurePopDensity.jpeg",
     width = 12, height = 5, dpi = 400, units = "in",
    device="jpeg")

############


# Main plot with jittered raw points and overlaid means ± SD
dens_summary <- FeatherAd_db %>%
  filter(!is.na(mean_density)) %>%  
  group_by(area_habitat, habitat, area) %>%
  summarise(
    an_mean_Tmean = first(an_mean_Tmean),
    mean_pop_density  = mean(mean_density),
    sd_density    = sd(mean_density),
    n = n(),
    .groups = "drop")

ggplot(dens_summary, aes(x = an_mean_Tmean, y = mean_pop_density)) +
  geom_pointrange(aes(
      ymin = mean_pop_density - sd_density,
      ymax = mean_pop_density + sd_density,
      fill = area,       
      shape = habitat), size = 1.5 ) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    colour = "tomato",
    linewidth = 0.7,
    alpha = 0.2) +
  scale_shape_manual(values = c("Urban" = 21, "Forest" = 22)) +
  scale_fill_brewer(palette = "Set2") +   # adjust fill colors
  # scale_colour_brewer(palette = "Set2") + 
  xlab("Annual mean temperature (°C)") +
  ylab("Barbule density") +
  theme_bw() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 18, hjust = 0.5),
    axis.text = element_text(size = 15, colour = "black"),
    axis.title = element_text(size = 18),
    plot.margin = unit(c(1, 1, 1, 1), "cm"))

### All variables in one panel
# Select the environmental columns you want
env_vars <- c("an_mean_Tmean", "an_se_Tmean", "an_mean_Tmin", "an_se_Tmin",
              "an_mean_Tmax", "an_se_Tmax", "win_mean_Tmean", "win_se_Tmean",
              "win_mean_Tmax", "win_se_Tmax", "win_mean_Tmin", "win_se_Tmin",
              "mou_mean_Tmean", "mou_se_Tmean", "mou_mean_Tmin",
              "mou_se_Tmin", "mou_mean_Tmax", "mou_se_Tmax")

# Calculate mean per population and SD for all the ENv variables
dens_summary_all <- lapply(env_vars, function(env) {
  FeatherAd_db %>%
    filter(!is.na(mean_length)) %>%
    group_by(area_habitat, habitat, area) %>%
    summarise(
      env_var = first(.data[[env]]),
      mean_pop_density = mean(mean_length),
      sd_density = sd(mean_length),
      n = n(),
      .groups = "drop"
    ) %>%
    mutate(env_name = env)
}) %>%
  bind_rows()

# Plot a panel
plots <- lapply(unique(dens_summary_all$env_name), function(env) {
  df <- dens_summary_all %>% filter(env_name == env)
  ggplot(df, aes(x = env_var, y = mean_pop_density)) +
    geom_pointrange(
      aes(
        ymin = mean_pop_density - sd_density,
        ymax = mean_pop_density + sd_density,
        fill = area,
        shape = habitat),size = 1.5, colour = "black") +
    geom_smooth(
      method = "lm",
      se = TRUE,
      colour = "tomato",
      linewidth = 0.7,
      alpha = 0.2) +
    scale_shape_manual(values = c("Urban" = 21, "Forest" = 22)) +
    scale_fill_brewer(palette = "Set2") +
    labs(x = env, y = "Barbule density") +
    theme_bw() +
    theme(
      legend.position = "none",
      plot.title = element_text(size = 18, hjust = 0.5),
      axis.text = element_text(size = 15, colour = "black"),
      axis.title = element_text(size = 18),
      plot.margin = unit(c(1, 1, 1, 1), "cm"))
})

panel <- wrap_plots(plots, ncol = 6)  # Adjust ncol for layout


ggsave(panel, file = "../Figures/Figure_EnvVar_Fstruct_Trial.jpeg",
     width = 30, height = 20, dpi = 400, units = "in",
    device="jpeg")

# ggsave(Figure1_May22, file = "../Figures/Figure1_Sep22.tiff",
#      width = 15, height = 12, dpi = 350, units = "in",
#     device="tiff")
```